---
title: "Homework #3"
author: "Suruchi Ahuja, Abbas Rizvi, Hoi Lam Tai, Jingchen Zhang"
date: "April 25, 2016"
output: pdf_document
---
```{r}
load("C:/Users/Suruchi Ahuja/Desktop/R prog/New folder/525 hw3/GeneMap.RData")
load("C:/Users/Suruchi Ahuja/Desktop/R prog/New folder/525 hw3/Array.RData")
load("C:/Users/Suruchi Ahuja/Desktop/R prog/New folder/525 hw3/Valid.RData")
```


Question - 4

```{r}
rm(list = ls())

# According to our biomarker challendge result, the top differencialy expressed features are 
#   145,705,235,63,754 and 954
# We return which gene set these features belong to
GeneMap[c(145,705,235,63,754,954)]
# 1  1 18 33 27 19
# these sets are also validated according to question 3(1,3,12,18,19,27,33,50)

# So we assume that the set 1,18,33,27,19 are differentially expressed, as fearture group "G1"
#   other features are "G2"


group=c()
for (i in 1:1000) {if (GeneMap[i] %in% c(1,18,33,27,19)) group[i]="G1" else group[i]="G2"}
GeneGrp=as.factor(group)

Values = function(i) t.test(ArrayData[i,c(1:3,7:13)],ArrayData[i,c(4:6,14:20)],var.equal=T)$p.value
ObsPvalue = mapply(Values, 1:1000)


plot(1:1000,ObsPvalue,col=GeneGrp)
abline(h=0.05,col="blue")


# factor object for significant genes
SigFactor=factor(ObsPvalue<=0.05, levels <- c(T,F))
levels(SigFactor)=c("Significant", "Non-Sig")

# creating a table
table(SigFactor,GeneGrp)

# test association of group with significance using Fisher's Exact Test
fisher.test(SigFactor,GeneGrp)


# Try other p-value's cut-off
# 0.01
SigFactor1=factor(ObsPvalue<=0.01, levels <- c(T,F))
levels(SigFactor1)=c("Significant", "Non-Sig")
table(SigFactor1,GeneGrp)
fisher.test(SigFactor1,GeneGrp)

# 0.1
SigFactor10=factor(ObsPvalue<=0.1, levels <- c(T,F))
levels(SigFactor10)=c("Significant", "Non-Sig")
table(SigFactor10,GeneGrp)
fisher.test(SigFactor10,GeneGrp)

```


Question - 5

Piano performs gene set analysis using various statistical methods, from different gene level statistics and a wide range of gene-set collections.Also, the Piano package contains functions for combining the results of multiple runs of gene set analyses.

```{r}
#To install this package :
#source("https://bioconductor.org/biocLite.R")
#biocLite("piano")

library(piano)

#Creating input 1: gene set collection 
genes2genesets = cbind(row.names(ArrayData), paste("GeneSet", GeneGrp))
myGeneSetCollection = loadGSC(genes2genesets)

#viewing all the genes in gene set G1
myGeneSetCollection$gsc[2]

# t-like method
#Creating input 2: gene-level statistic (method 1)
#create a binary response
Y=c(rep(0,3),rep(1,3),rep(0,7),rep(1,7))
icor=function(i) cor(ArrayData[i,Y==0],ArrayData[i,Y==1])
myStats=mapply(icor,1:1000)
names(myStats)=rownames(ArrayData)
print(round(myStats,3)[1:10])

GSA = runGSA(myStats, gsc=myGeneSetCollection, adjMethod="bonferroni")

GSA

Results=GSAsummaryTable(GSA)
#viewing table of results
head(Results)
#identifying the labels so we can extract the resuls of interest
colnames(Results)

Results["p (non-dir.)"]
Results["p adj (non-dir.)"]

#################
# p-like method
#Creating input 2: gene-level statistic (method 2)
myStats=ObsPvalue
names(myStats)=rownames(ArrayData)
print(round(myStats,5)[1:7])

gsaF = runGSA(geneLevelStats=myStats,
              directions=myStats,
              geneSetStat = "fisher",
              gsc=myGeneSetCollection,
              adjMethod="bonferroni")


#viewing the specifics of the GSA analysis
gsaF
#making a table of the results
Rest=GSAsummaryTable(gsaF)
table(Rest)
#extracting the possible information available
colnames(Rest)
#selecting the results of interest
Rest["p adj (non-dir.)"]
Rest["p adj (mix.dir.up)"]


```

Question - 6 

Performing the GSEA analysis on the Gene Map data.
GSEA - Gene Set Enrichment Analysis is a computational method that determines whether an a priori defined set of genes shows statistically significant, concordant differences between two biological states (e.g. phenotypes).

```{r}

# Using the GeneMap.RData, perform a GSEA analysis of the type that we discussed in lecture.

SmplGrp=c(rep(0,3),rep(1,3),rep(0,7),rep(1,7)) # 0=Group A, 1=Group B

# We could/(should?) use t-test values instead of correlation but
# for illustrative purposes, let's use correlation

MyCor=function(i) cor(SmplGrp,ArrayData[i,])
RhoHats=mapply(MyCor,1:1000)

# aside, let's compare t-test p-vals and cor estimates
plot(ObsPvalue,RhoHats)

# let's look at corr values
plot(1:1000,RhoHats,col=GeneMap)

# let's sort the values in descending order
oDX=order(RhoHats,decreasing=T)
plot(1:1000,RhoHats[oDX],col=GeneMap[oDX],xlab="ordered gene list")

# let's make this look like a component in the GSEA plots that we have seen:
plot(1:1000,RhoHats[oDX],col=GeneMap[oDX],xlab="ordered gene list",ylab=expression(hat(rho)),pch=" ")
abline(h=0)
for(i in 1000:1) lines(c(i,i),c(0,RhoHats[oDX[i]]),col=GeneMap[oDX[i]])

ranks=1001-rank(RhoHats) # rank such that highest corr has rank 1

# put vertical lines for group A
abline(v=ranks[1:10])

```


